\documentclass[10pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}
\usepackage{geometry}
\geometry{margin=1.8cm}
\usepackage{enumitem}
\usepackage{booktabs}
\usepackage{fancyhdr}
\usepackage{lastpage}
\usepackage{tcolorbox}
\usepackage{hyperref}

\pagestyle{fancy}
\fancyhf{}
\rfoot{\thepage/\pageref{LastPage}}
\lhead{\small TOEFL ITP 2026 --- Day 8}
\rhead{\small MPrometeo}
\renewcommand{\headrulewidth}{0.4pt}
\setlength{\parindent}{0pt}
\setlength{\parskip}{5pt}

\begin{document}

\begin{center}
{\LARGE \textbf{DAY 8 --- Friday Mar 6}}\\[4pt]
{\large Listening Patterns + Passive Voice}\\[2pt]
{\normalsize Theme: Trump Administration --- Wins \& Failures 2025--26}\\[2pt]
\rule{\textwidth}{0.8pt}
\end{center}

% ============================================================
\section*{Grammar: Passive Voice}
% ============================================================

\subsection*{Why passive matters for TOEFL}

Academic and political English \emph{loves} passive voice. It
shifts focus from who did something to what was done --- essential
for objective, formal writing. TOEFL Structure tests this constantly.

\subsection*{Formation}

\begin{center}
\begin{tabular}{lll}
\toprule
\textbf{Tense} & \textbf{Active} & \textbf{Passive} \\
\midrule
Simple Present & Trump signs the bill & The bill \textbf{is signed} \\
Simple Past & Congress passed the law & The law \textbf{was passed} \\
Present Perfect & They have raised tariffs & Tariffs \textbf{have been raised} \\
Past Perfect & He had fired the director & The director \textbf{had been fired} \\
Modal & They will implement cuts & Cuts \textbf{will be implemented} \\
Modal Perfect & They should have warned & People \textbf{should have been warned} \\
\bottomrule
\end{tabular}
\end{center}

\subsection*{Key Rules to Memorize}

\begin{enumerate}[nosep]
\item Passive = \texttt{be + past participle} (ALWAYS)
\item Only \textbf{transitive} verbs can be passive (they need an object)
\item ``By + agent'' is optional: ``The law was passed (by Congress).''
\item \textbf{TOEFL trap:} missing ``be'' $\to$ ``The bill signed'' is WRONG
\item \textbf{TOEFL trap:} wrong participle $\to$ ``was sign'' is WRONG (must be ``signed'')
\end{enumerate}

\subsection*{Passive with Reporting Verbs (very common in TOEFL)}

\begin{itemize}[nosep]
\item It \textbf{is believed} that tariffs reduce imports.
\item The policy \textbf{is considered} controversial.
\item The president \textbf{was reported} to have signed the order.
\item It \textbf{is estimated} that 200,000 jobs were affected.
\end{itemize}

% ============================================================
\section*{Reading Passage 1: Trump's Policy Wins ($\sim$230 words)}
% ============================================================

\begin{tcolorbox}[colback=white, colframe=black]
\small
The second Trump administration entered office in January 2025 with
an aggressive economic agenda. Within its first year, several
measurable outcomes were achieved.

The ``One Big Beautiful Bill'' made the 2017 tax cuts for working
families permanent, created new ``Trump Accounts'' savings vehicles,
and eliminated taxes on tips --- a popular measure among service
workers. The administration reported nearly 4\% wage growth and a
3\% GDP expansion in the first half of 2025, driven partly by strong
consumer spending.

On trade, Trump implemented a universal 10\% tariff on most imports
and negotiated a significant framework deal with China. Beijing
agreed to suspend retaliatory tariffs on American agricultural
products, resume soybean imports, eliminate export controls on rare
earth minerals, and cooperate on fentanyl exports. The Council on
Foreign Relations noted that while the deals were ``transactional
and temporary,'' they represented concrete results.

The Department of Government Efficiency (DOGE), led by Elon Musk,
claimed over \$160 billion in federal savings through workforce
reductions and spending cuts, though the actual net savings remain
disputed.

However, economists warned that short-term metrics can be misleading.
Stock markets experienced significant volatility after tariff
announcements, and independent analysts cautioned that higher import
costs would eventually reach consumers. The administration's
supporters argue these are necessary growing pains; critics call
them reckless gambles with the world's largest economy.
\end{tcolorbox}

\subsection*{Questions}

\begin{enumerate}[label=\textbf{\arabic*.}]
\item The main idea of the passage is:
\begin{enumerate}[label=(\Alph*)]
\item Trump's policies had no effect
\item The administration achieved measurable economic results but
with significant debate about sustainability
\item The economy collapsed under Trump
\item China won the trade war
\end{enumerate}

\item ``Transactional and temporary'' suggests the trade deals:
\begin{enumerate}[label=(\Alph*)]
\item Are permanent treaties
\item Could change at any time and were based on mutual leverage
\item Were ignored by both sides
\item Only benefited China
\end{enumerate}

\item The word ``disputed'' is closest in meaning to:
\begin{enumerate}[label=(\Alph*)]
\item confirmed \item celebrated \item debated/challenged \item ignored
\end{enumerate}

\item What concern do independent economists raise?
\begin{enumerate}[label=(\Alph*)]
\item Wages are too high
\item Short-term gains may hide long-term costs like inflation
\item The GDP is too low
\item Trade deals are too permanent
\end{enumerate}

\item It can be inferred that DOGE's savings claims:
\begin{enumerate}[label=(\Alph*)]
\item Are universally accepted as accurate
\item May be exaggerated or offset by hidden costs
\item Saved exactly \$2 trillion
\item Had no impact on federal employees
\end{enumerate}
\end{enumerate}

\textbf{Answers:} 1-B, 2-B, 3-C, 4-B, 5-B

% ============================================================
\section*{Reading Passage 2: Trump's Failures ($\sim$230 words)}
% ============================================================

\begin{tcolorbox}[colback=white, colframe=black]
\small
For every claimed achievement, critics point to a corresponding
failure. The Trump administration's second term has been described
by Mediaite as a ``chaos machine'' --- a deliberate strategy of
overwhelming the public with simultaneous scandals, making sustained
accountability nearly impossible.

On his first day back in office, Trump pardoned hundreds of January
6th defendants, a decision that even some Republican lawmakers
called a dangerous signal about the rule of law. The firing of
federal inspectors general --- independent watchdogs designed to
prevent government abuse --- removed key oversight mechanisms. Over
200 executive orders were issued in the first year alone, many
bypassing Congress entirely.

Immigration policy drew international condemnation. Mass
deportations to third countries with limited judicial oversight,
military deployment in civilian areas, and the separation of
families at the border were criticized by Human Rights Watch as
``pervasive attacks on rights.'' Federal courts repeatedly struck
down several of these measures as unconstitutional.

The administration's relationship with foreign governments also
raised ethical concerns. Transactions involving cryptocurrency,
Gulf-state aircraft deals, and the seizure of a Venezuelan oil
tanker were characterized by The New Republic as ``the most blatant
instances of presidential self-interest in modern American history.''

Perhaps most significantly, approval ratings declined steadily
throughout 2025, suggesting that even supporters were growing weary
of what critics call governance through permanent crisis.
\end{tcolorbox}

\subsection*{Questions}

\begin{enumerate}[label=\textbf{\arabic*.}, start=6]
\item The ``chaos machine'' strategy refers to:
\begin{enumerate}[label=(\Alph*)]
\item Efficient government operations
\item Overwhelming the public with simultaneous scandals to
prevent accountability
\item A new government department
\item Military strategy
\end{enumerate}

\item ``Pervasive'' is closest in meaning to:
\begin{enumerate}[label=(\Alph*)]
\item minor \item widespread and persistent
\item temporary \item invisible
\end{enumerate}

\item What happened to the inspectors general?
\begin{enumerate}[label=(\Alph*)]
\item They were promoted
\item They were fired, removing independent oversight
\item They resigned voluntarily
\item They supported all policies
\end{enumerate}

\item The passage suggests the Venezuelan oil tanker seizure was:
\begin{enumerate}[label=(\Alph*)]
\item A routine military operation
\item An act driven by presidential self-interest
\item Supported by all parties
\item Requested by Venezuela
\end{enumerate}

\item The declining approval ratings suggest:
\begin{enumerate}[label=(\Alph*)]
\item Universal support for the president
\item Growing public fatigue with governance through crisis
\item The polls were inaccurate
\item Approval was rising
\end{enumerate}
\end{enumerate}

\textbf{Answers:} 6-B, 7-B, 8-B, 9-B, 10-B

% ============================================================
\section*{Passive Voice Exercises (10 questions)}
% ============================================================

\textit{All themed around Trump-era politics. Choose correctly.}

\begin{enumerate}[label=\textbf{\arabic*.}, leftmargin=*]

\item The tax bill \rule{2cm}{0.4pt} by Congress in 2025.
\begin{enumerate}[label=(\Alph*)]
\item passed \item was passed \item has passing \item passing
\end{enumerate}

\item Over 200 executive orders \rule{2cm}{0.4pt} in the first year.
\begin{enumerate}[label=(\Alph*)]
\item issued \item were issued \item have issuing \item was issued
\end{enumerate}

\item The tariffs \rule{2cm}{0.4pt} to reduce the trade deficit.
\begin{enumerate}[label=(\Alph*)]
\item are designed \item designed \item is designing \item designs
\end{enumerate}

\item It \rule{2cm}{0.4pt} that the policy affected 200,000 workers.
\begin{enumerate}[label=(\Alph*)]
\item estimates \item is estimated \item estimated \item estimating
\end{enumerate}

\item The inspectors general \rule{2cm}{0.4pt} before they could
complete their investigations.
\begin{enumerate}[label=(\Alph*)]
\item fired \item were fired \item have firing \item was fired
\end{enumerate}

\item \textit{Find error:}
The trade deal (A)\textbf{between} the US and China
(B)\textbf{was announce} (C)\textbf{after} months of
(D)\textbf{negotiations}.

\item Several immigration measures \rule{2cm}{0.4pt} unconstitutional
by federal courts.
\begin{enumerate}[label=(\Alph*)]
\item declared \item were declared \item declaring \item has declared
\end{enumerate}

\item \textit{Find error:}
The pardons (A)\textbf{were} (B)\textbf{criticize} by
(C)\textbf{both} Democrats and (D)\textbf{some} Republicans.

\item The department \rule{2cm}{0.4pt} by Elon Musk since its
creation in early 2025.
\begin{enumerate}[label=(\Alph*)]
\item has been led \item was leading \item led \item has leading
\end{enumerate}

\item Had the oversight mechanisms not \rule{2cm}{0.4pt}, the
abuses might have been caught earlier.
\begin{enumerate}[label=(\Alph*)]
\item removed \item been removed \item removing \item be removed
\end{enumerate}
\end{enumerate}

\subsection*{Answer Key}

\begin{center}
\begin{tabular}{cll}
\toprule
\# & Ans & Explanation \\
\midrule
1 & B & Simple past passive: was + past participle \\
2 & B & Plural subject ``orders'' $\to$ ``were issued'' \\
3 & A & Present passive: ``are designed'' \\
4 & B & Reporting passive: ``It is estimated that...'' \\
5 & B & Plural ``inspectors'' $\to$ ``were fired'' \\
6 & B & ``Was announce'' $\to$ ``was announced'' (need pp) \\
7 & B & Past passive: ``were declared'' \\
8 & B & ``Criticize'' $\to$ ``criticized'' (need pp) \\
9 & A & Present perfect passive: ``has been led'' \\
10 & B & Past perfect passive: ``had not been removed'' \\
\bottomrule
\end{tabular}
\end{center}

% ============================================================
\section*{Listening: Idioms \& Expressions (30 to memorize)}
% ============================================================

These appear constantly in TOEFL ITP listening:

\begin{center}
\small
\begin{tabular}{lll}
\toprule
\textbf{Expression} & \textbf{Meaning} & \textbf{Context} \\
\midrule
across the board & affecting everything & Tariffs across the board \\
a far cry from & very different & A far cry from promises \\
behind closed doors & in secret & Deals behind closed doors \\
bite off more than you can chew & take on too much & DOGE's goals \\
by and large & generally speaking & By and large, the economy grew \\
catch-22 & no-win situation & Tariffs: raise or lower? \\
come under fire & be criticized & Policy came under fire \\
cut both ways & have good and bad effects & Trade cuts both ways \\
dead in the water & completely stalled & The bill was dead in the water \\
down the road & in the future & Consequences down the road \\
get the ball rolling & start something & Get reforms rolling \\
in hot water & in trouble & The president is in hot water \\
in the long run & eventually & Tariffs hurt in the long run \\
jump the gun & act too soon & Jumped the gun on tariffs \\
on the fence & undecided & Voters on the fence \\
\bottomrule
\end{tabular}
\end{center}

% ============================================================
\section*{Shadowing (Gemini Live --- 15 min)}
% ============================================================

\begin{quote}\small
``The second Trump administration achieved measurable economic
results, including nearly four percent wage growth and a three
percent GDP expansion. However, economists warned that short-term
metrics can be misleading. Stock markets experienced significant
volatility, and independent analysts cautioned that higher import
costs would eventually reach consumers.''
\end{quote}

\textbf{Targets:} \textbf{administration} (ad-min-is-TRAY-shun),
\textbf{measurable} (MEZH-ur-a-bul), \textbf{volatility}
(vol-a-TIL-i-tee), \textbf{analysts} (AN-a-lists).

\end{document}